Allocation Removal by Partial Evaluation in a Tracing JIT

### 2011-01-11 talk/pepm2011/presentation/talk.tex
continue working on the talk. will have to do things differently, the diagrams are nice but take way too much space.

\begin{frame}
  \frametitle{Optimizing the Heap Operations in a Trace}
  \begin{itemize}
      \item Contribution of our paper
      \item A simple, efficient and effective optimization of heap operations in a trace
      \item using online partial evaluation
  \end{itemize}
\end{frame}

@@

@@

>>>>

\begin{frame}
  \frametitle{Optimizing the Heap Operations in a Trace}
  \begin{itemize}
      \item Contribution of our paper
      \item A simple, efficient and effective optimization of heap operations in a trace
      \item using online partial evaluation
      \item fully implemented an in use in large-scale interpreters
  \end{itemize}
  \pause
  \begin{block}{Ingredients}
      \begin{itemize}
          \item a slightly simplified model for objects on the heap
          \item operational semantics of trace operations that manipulate the heap
          \item optimization rules for those operations
      \end{itemize}
  \end{block}
\end{frame}

@@

\begin{frame}
  \frametitle{Operations Manipulating Heap Objects}
  \begin{itemize}
      \item \texttt{v = new(T)} makes a new object
      \item \texttt{get(w, F)} reads a field out of an object
      \item \texttt{u = set(v, F, w)} writes a field of an object
      \item \texttt{guard(v, T)} checks the type of an object
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Optimization by Online Partial Evaluation}
  \begin{itemize}
      \item Trace is optimized using online partial evaluation
      \item part of the runtime heap is modelled in the \emph{static heap}
      \item static heap contains objects that are allocated within the trace
      \item (as opposed to before the trace is executed)
      \pause
      \item operations acting on the static heap can be removed
      \item all others need to be residualized
      \item all fairly straightforward
  \end{itemize}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing New}
  \includegraphics[scale=0.8]{figures/opt_new1}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing New}
  \includegraphics[scale=0.8]{figures/opt_new2}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Get}
  \includegraphics[scale=0.8]{figures/opt_get1}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Get}
  \includegraphics[scale=0.8]{figures/opt_get2}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Get}
  \includegraphics[scale=0.8]{figures/opt_get3}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Get}
  \includegraphics[scale=0.8]{figures/opt_get4}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Guard}
  \includegraphics[scale=0.8]{figures/opt_guard1}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Guard}
  \includegraphics[scale=0.8]{figures/opt_guard2}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Guard}
  \includegraphics[scale=0.8]{figures/opt_guard3}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Guard}
  \includegraphics[scale=0.8]{figures/opt_guard4}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Set}
  \includegraphics[scale=0.8]{figures/opt_set1}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Set}
  \includegraphics[scale=0.8]{figures/opt_set2}
\end{frame}

\begin{frame}
  \frametitle{Lifting}
  Problem: What happens if we write a static object into a dynamic one?
  \pause
  \begin{itemize}
      \item lose track of the static object because of possibility of aliasing
      \item need to \emph{lift} the static object
      \item lifting produces operations that recreate the static object
  \end{itemize}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Set}
  \includegraphics[scale=0.8]{figures/opt_set_dynamic1}
\end{frame}

\begin{frame}[plain]
  \frametitle{Optimizing Set}
  \includegraphics[scale=0.8]{figures/opt_set_dynamic2}
\end{frame}

\begin{frame}
  \frametitle{Benchmark Results}
  \begin{itemize}
      \item to evaluate the optimization we used PyPy's Python interpreter with real-world programs
      \item optimization can remove
      \begin{itemize}
          \item 70\% of all \texttt{new} operations
          \item 14\% of all \texttt{get/set} operations
          \item 93\% of all \texttt{guard} operations
      \pause
      \item Timings improve by a factor between 1.2 and 6.95
      \item outperforming standard Python on all benchmarks but one
      \pause
      \item more details in the paper
      \end{itemize}
  \end{itemize}
\end{frame}
